When are Fuzzy Extractors Possible?

### 2015-02-11 trunk/info-bounds/FuzzyExtPossibility-crypto.tex

\title{When are Fuzzy Extractors Possible?}
\blind{
\author{
Benjamin Fuller\footnote{Email: {\tt bfuller@cs.bu.edu}.  Boston
  University and MIT Lincoln Laboratory.}  \and Leonid
Reyzin\footnote{Email: {\tt reyzin@cs.bu.edu}.  Boston University.}
\and Adam Smith\footnote{Email: {\tt asmith@cse.psu.edu}.
  Pennsylvania State University. This work performed while visiting Boston
  University's Hariri Institute for Computing.}}
}
\begin{document}
\maketitle

\begin{abstract}
Fuzzy extractors (Dodis et al., Eurocrypt 2004) convert repeated noisy readings of a high-entropy secret into the same uniformly distributed key. A minimum condition for the security of the key is the hardness of guessing a value that is similar to the secret, because the fuzzy extractor converts such a guess to the key.
We define \emph{fuzzy min-entropy} to quantify this property of a noisy source of secrets.  Fuzzy min-entropy measures the success of the adversary when provided with \emph{only} the functionality of the fuzzy extractor, that is, the \emph{ideal} security possible from a noisy distribution.

High fuzzy min-entropy is necessary for the existence of a fuzzy extractor; moreover, there is evidence that it may be sufficient when only computational security is required. Nevertheless, information-theoretic fuzzy extractors are not known for many practically relevant sources of high fuzzy min-entropy. In this work, we ask whether this challenge is inherent.  We show the following:
\begin{itemize}
\item \textbf{Sufficiency of fuzzy min-entropy:}  We construct a fuzzy extractor for each physical source with fuzzy min-entropy.  This result justifies fuzzy min-entropy as the right measure of the quality of a noisy distribution for key derivation.  Previous results considered the entropy and error tolerance of a separately and had not secured many sources with fuzzy min-entropy.  
The constructed fuzzy extractor must \emph{precisely know} the distribution of the noisy source.
\item \textbf{Cost of distributional uncertainty:}  High entropy sources can never be fully observed and must be modeled.  An adversary may have a better model than the fuzzy extractor designer.  To compensate, fuzzy extractors are typically designed to work for families of physical sources. We build a family of high fuzzy min-entropy sources which no fuzzy extractor can secure.  This result emphasizes the importance of high fidelity models of high entropy sources.
\end{itemize}
We provide similar but stronger results for secure sketches, whose goal is not to derive a consistent key, but to recover a consistent reading of the secret.  Our results are information-theoretic, but we discuss implications for the computational setting.

@@

\paragraph{Fuzzy Min-Entropy}
We introduce a new entropy notion that precisely measures how hard it is for the adversary to guess a value within distance $t$ of the original reading $w$, thus subverting the security of $\key$ by running $\rep$. Suppose $w$ is sampled from a distribution $W$.   To have the maximum chance that $w'$ is within distance $t$ of $w$, the adversary would want to maximize the total probability mass of $W$ within the ball $B_t(w')$ of radius $t$ around $w'$.
We  therefore define \emph{fuzzy min-entropy} $\Hfuzz(W) \eqdef -\log \max_{w'} \Pr[W\in B_t(w')]$.  Observe that this quantity can be bounded in terms of min-entropy: $\Hoo(W) \ge \Hfuzz(W) \ge \Hoo(W)-\log |B_t|$.

Fuzzy min-entropy measures the quality of a distribution when provided oracle access to the functionality of a fuzzy extractor.  Fuzzy min-entropy measures the ideal security of a noisy source of entropy when used for key derivation, similar to distributional notions of program obfuscation~\cite{barak2001possibility,DBLP:conf/stoc/DodisS05,DBLP:conf/tcc/DodisS05}---we discuss the connection in \secref{sec:related settings}.  Superlogarithmic fuzzy min-entropy is \emph{necessary} for nontrivial key extraction~(\propref{prop:fuzz necessary} formalizes the above intuition).

However, existing constructions do not measure their security in terms of fuzzy min-entropy; instead, their security is shown to be  $\Hoo(W)$ minus some loss that is at least $\log |B_t|$ due to error-tolerance. Since $\Hoo(W)-\log |B_t| \le \Hfuzz(W)$, it is natural to ask whether this loss is necessary. This question is particularly relevant when the gap between the two sides of the inequality is high.\footnote{For nearly uniform distributions, $\Hfuzz(W) \approx \Hoo(W)- \log |B_t|$.  In this setting, standard coding based constructions of fuzzy extractors (instantiated with optimum codes) yield keys of size approximately $\Hfuzz(W)$.}  As an example, iris scans appear to have significant $\Hfuzz(W)$ (because iris scans for different people appear to be well-spread in the metric space~\cite{daugman2006probing}) but negative $\Hoo(W) -\log |B_t|$ \cite[Section 5]{blanton2009biometric}. We therefore ask: \emph{is fuzzy min-entropy sufficient for fuzzy extraction?} There is evidence that it may be when the security requirement is computational rather than information-theoretic---see \secref{sec:related settings}. We provide an answer in two settings.

\paragraph{Sufficiency of $\Hfuzz(W)$ for a Precisely Known Distribution}
Ideally, a fuzzy extractor has \emph{precise knowledge} of the probability distribution function of $W$.  We call this  the \emph{precise knowledge} setting.  In this setting, we show that for every source $W$ with superlogarithmic $\Hfuzz(W)$, it is possible to construct a fuzzy extractor with a superlogarithmic length $\key$ (\corref{cor:extension to fuzz ext}). Our construction crucially utilizes the probability distribution function of $W$ and, in particular, is not polynomial time.  This result shows that $\Hfuzz(W)$ is a necessary and sufficient condition for building a fuzzy extractor of a distribution $W$.

@@

\paragraph{The Challenge of Precisely Distributional Knowledge}
Assuming \emph{precise knowledge} of a distribution $W$ may be unrealistic. Indeed, high-entropy distributions can never be fully observed directly and must therefore be modeled. It is imprudent to assume that the designer's model of a distribution is completely accurate---the adversary, with greater resources, would likely be able to build a better model. Therefore, fuzzy extractor designs cannot usually be tailored to one  particular source. Existing designs work for a family of sources (for example, all sources of min-entropy at least $m$ with at most $t$ errors). Thus, the design is fixed with a partially known distribution, and the adversary may know more about the distribution than the designer of the fuzzy extractor.  We call this the \emph{distributional uncertainty} setting.

@@

We show comparable, but stronger, results for secure sketches.  Namely, we show in \corref{cor:extension to fuzz ext} that secure sketches are possible if the distribution $W$ is precisely known. (In fact, we obtain our fuzzy extractors for the case of a known distribution from this result by applying a randomness extractor.)

>>>>

\title{When are Fuzzy Extractors Possible?}
\blind{
\author{
Benjamin Fuller\footnote{Email: {\tt bfuller@cs.bu.edu}.  Boston
  University and MIT Lincoln Laboratory.}  \and Leonid
Reyzin\footnote{Email: {\tt reyzin@cs.bu.edu}.  Boston University.}
\and Adam Smith\footnote{Email: {\tt asmith@cse.psu.edu}.
  Pennsylvania State University. This work performed while visiting Boston
  University's Hariri Institute for Computing.}}
}
\author{}\institute{}
\begin{document}
\maketitle

\begin{abstract}
Fuzzy extractors (Dodis et al., Eurocrypt 2004) convert repeated noisy
readings of a high-entropy secret into the same uniformly distributed
key. A minimum condition for the security of the key is the hardness
of guessing a value that is similar to the secret, because the fuzzy
extractor converts such a guess to the key.

We define \emph{fuzzy min-entropy} to quantify this property of a noisy source of secrets.  Fuzzy min-entropy measures the success of the adversary when provided with \emph{only} the functionality of the fuzzy extractor, that is, the \emph{ideal} security possible from a noisy distribution.  
High fuzzy min-entropy is necessary for the existence of a fuzzy
extractor. 
% ; moreover, there is evidence that it may be sufficient when only computational security is required.

We ask: \emph{is high fuzzy min-entropy a
  sufficient condition for noise-tolerant extraction?}
If only computational security is required, recent progress on program
obfuscation gives evidence that fuzzy
min-entropy is indeed sufficient. In contrast,
information-theoretic fuzzy extractors are not known for many
practically relevant sources of high fuzzy min-entropy.

In this paper,
show that fuzzy min-entropy is also sufficient for
information-theoretically secure fuzzy extraction. For every source
distribution $W$ for which security is possible in an ideal world with
access only to the extractor functionality, we give a secure fuzzy
extractor.

Our construction relies on the extractor knowing the precise
distribution of the source $W$.  A more ambitious goal is to design a
single extractor that works for all possible sources.  We show that
this more ambitious goal is impossible: we give a family of sources
with high fuzzy min-entropy for which no single fuzzy extractor is
secure.  This result emphasizes the importance of accurate models
of high entropy sources.

% \begin{itemize}
% \item \textbf{Sufficiency of fuzzy min-entropy:} We show that for each
%   source with nontrivial (that is, $\omega(\log n)$) fuzzy min-entropy,
%   there is an infor\-mation-theoretic fuzzy extractor that reliably
%   extracts a key of length roughly the fuzzy min-entropy.  This result
%   justifies fuzzy min-entropy as the ``right'' measure of the quality
%   of a noisy distribution for key derivation. Previous results
%   considered the entropy and error tolerance of a source separately;
%   consequently, for many sources with high fuzzy min-entropy, existing
%   constructions did not lead to secure fuzzy extractors.

%   Our construction works for every source, but requires
%   \emph{precise knowledge} of the distribution of the source.

% \item \textbf{Cost of distributional uncertainty:} A more ambitious
%   goal is to design a single extractor that works for all possible
%   sources.

% High entropy
%   sources can never be fully observed by sampling, and so some
%   modeling assumptions must be used.

% Moreover,  An
%   adversary may have a better model than the fuzzy extractor designer.
%   To compensate, fuzzy extractors are typically designed to work for
%   families of physical sources. We build a family of high fuzzy
%   min-entropy sources which no fuzzy extractor can secure.  This
%   result emphasizes the importance of high fidelity models of high
%   entropy sources.
% \end{itemize}
% We provide similar but stronger results for secure sketches, whose goal is not to derive a consistent key, but to recover a consistent reading of the secret.  Our results are information-theoretic, but we discuss implications for the computational setting.

@@

\paragraph{Fuzzy Min-Entropy}
We introduce a new entropy notion that precisely measures how hard it
is for the adversary to guess a value within distance $t$ of the
original reading $w$. 
Suppose $w$ is sampled from a distribution $W$.   To have the maximum chance that $w'$ is within distance $t$ of $w$, the adversary would want to maximize the total probability mass of $W$ within the ball $B_t(w')$ of radius $t$ around $w'$.
We  therefore define \emph{fuzzy min-entropy} $\Hfuzz(W) \eqdef -\log \max_{w'} \Pr[W\in B_t(w')]$.  Observe that this quantity can be bounded in terms of min-entropy: $\Hoo(W) \ge \Hfuzz(W) \ge \Hoo(W)-\log |B_t|$.

% Fuzzy min-entropy measures the quality of a distribution when provided oracle access to the functionality of a fuzzy extractor.  
Fuzzy min-entropy measures the ideal security of a noisy source of
entropy when used for key derivation, similar to distributional
notions of program obfuscation
(e.g.,~\cite{barak2001possibility,DBLP:conf/stoc/DodisS05,DBLP:conf/tcc/DodisS05};
see \secref{sec:related settings} for discussion).  Superlogarithmic fuzzy min-entropy is \emph{necessary} for nontrivial key extraction~(\propref{prop:fuzz necessary}).

However, existing constructions do not measure their security in terms
of fuzzy min-entropy; instead, their security is shown to be
$\Hoo(W)$ minus some loss, for error-tolerance, that is at least $\log |B_t|$. Since $\Hoo(W)-\log |B_t| \le \Hfuzz(W)$, it is natural to ask whether this loss is necessary. This question is particularly relevant when the gap between the two sides of the inequality is high.\footnote{For nearly uniform distributions, $\Hfuzz(W) \approx \Hoo(W)- \log |B_t|$.  In this setting, standard coding based constructions of fuzzy extractors (instantiated with optimum codes) yield keys of size approximately $\Hfuzz(W)$.}  As an example, iris scans appear to have significant $\Hfuzz(W)$ (because iris scans for different people appear to be well-spread in the metric space~\cite{daugman2006probing}) but negative $\Hoo(W) -\log |B_t|$ \cite[Section 5]{blanton2009biometric}. We therefore ask: \emph{is fuzzy min-entropy sufficient for fuzzy extraction?} There is evidence that it may be when the security requirement is computational rather than information-theoretic---see \secref{sec:related settings}. We provide an answer in two settings.

\paragraph{Sufficiency of $\Hfuzz(W)$ for a Precisely Known Distribution}
Ideally, a fuzzy extractor has \emph{precise knowledge} of the
probability distribution function of $W$.  We call this  the
\emph{precise knowledge} setting.  In this setting, we show that for
every source $W$ with superlogarithmic $\Hfuzz(W)$, it is possible to
construct a fuzzy extractor with a superlogarithmic length $\key$
(\corref{cor:extension to fuzz ext}). Our construction crucially
utilizes the probability distribution function of $W$ and, in
particular, is not polynomial time.  This result shows that
$\Hfuzz(W)$ is a necessary and sufficient condition for building a
fuzzy extractor for a given distribution $W$.

@@

\paragraph{The Challenge of Precise Distributional Knowledge}
Assuming precise knowledge of a distribution $W$ may be unrealistic. Indeed, high-entropy distributions can never be fully observed directly and must therefore be modeled. It is imprudent to assume that the designer's model of a distribution is completely accurate---the adversary, with greater resources, would likely be able to build a better model. Therefore, fuzzy extractor designs cannot usually be tailored to one  particular source. Existing designs work for a family of sources (for example, all sources of min-entropy at least $m$ with at most $t$ errors). Thus, the design is fixed with a partially known distribution, and the adversary may know more about the distribution than the designer of the fuzzy extractor.  We call this the \emph{distributional uncertainty} setting.

@@

For secure sketches, we show results that are similar to but
stronger than our results for fuzzy extractor.  Namely, we show in \corref{cor:extension to fuzz ext} that secure sketches are possible if the distribution $W$ is precisely known. (In fact, we obtain our fuzzy extractors for the case of a known distribution from this result by applying a randomness extractor.)
